---
title: "p8105_hw2_hh3043"
author: "Haitian Huang"
date: "2023-10-02"
output: html_document
---
```{r}
library(tidyverse)
library(readxl)
```

problem 1

First,clean the pols_month dataset.
```{r}

pols_month = 
  read_csv('.\\fivethirtyeight_datasets\\pols-month.csv')|>
  separate(mon,into = c('year','month','day'),convert=TRUE) |>
  mutate(month=month.name[month])|>
  mutate(
    president = case_match(prez_gop, 0 ~ "dem",
                         1 ~ "gop", 
                         2 ~ "gop"))|>
  select(year,month,everything(),-day,-starts_with('prez_'))





```
Then, clean the snp dataset.

```{r}
snp = read_csv('.\\fivethirtyeight_datasets\\snp.csv')|>
  separate(date,into = c('day','month','year'),convert = TRUE) |>
  mutate(
    year = if_else(year>23,year+1900,year+2000),
    month = month.name[month]
  )
```
Third,clean the unemployment data.
```{r}
unemployment = read_csv('.\\fivethirtyeight_datasets\\unemployment.csv')|> 
  rename(year = Year) |>
  pivot_longer(
    Jan:Dec, 
    names_to = "month",
    values_to = "unemployment"
  ) |>
  mutate(
    month = month.name[match(month,month.abb)]
  )
```
Finally,join all three data sets.

```{r}
data_538 = left_join(pols_month,snp)|>
  left_join(x=_,unemployment)
```

problem 2

Clean the Mr.Trash data set and dcalculate the homes powered.
```{r}



Mr_Trash = read_excel('202309 Trash Wheel Collection Data.xlsx',sheet=1,range = cell_cols('A:N'))|>
  janitor::clean_names() |>
  rename(homes_powered_approximate = homes_powered)|>
  drop_na('dumpster')|>
  mutate(
    year = as.numeric(year),
    homes_powered = weight_tons*500/30,
    category = 'Mr_Trash')
```


```{r}
Professor = read_excel('202309 Trash Wheel Collection Data.xlsx',sheet=2,range = cell_cols('A:M'))|>
  janitor::clean_names() |>
  rename(homes_powered_approximate = homes_powered)|>
  drop_na('dumpster')|>
  mutate(
    homes_powered = weight_tons*500/30,
    category = 'Professor')


```

```{r}
Gwynnda = read_excel('202309 Trash Wheel Collection Data.xlsx',sheet=4,range = cell_cols('A:L'))|>
  janitor::clean_names() |>
  rename(homes_powered_approximate = homes_powered)|>
  drop_na('dumpster')|>
  mutate(
    homes_powered = weight_tons*500/30,
    category = 'Gwynnda')
```



```{r}
Trash_Wheel = bind_rows(Mr_Trash,Professor)|>
  bind_rows(x=_,Gwynnda)|>
  arrange(dumpster)
```
These trash wheel data are collected are about the amount and categories of the trash the water wheel vessels removed.For `Mr.Trash` dataset,there are `r nrow(Mr_Trash)` observations and `r ncol(Mr_Trash)` variables.For `Professor` dataset, there are 
`r nrow(Professor)` observations and `r ncol(Professor)` variables. For `Gwynnda` dataset,there are `r nrow(Gwynnda)` observations and `r ncol(Gwynnda)` variables. The `weight_tons` variable is the weight of trash remove of the trash wheel in one day.The `homes_powered` variable is how many homes can be supported by the electricity transferred by the trash collected on that day.The totoal weight of trash collected by Professor Trash_Wheel is `r sum(pull(Professor,weight_tons))`.The total number of cigarette butts collected by Gwynnda Trash Wheel is `r sum(pull(Gwynnda,cigarette_butts))`.


Problem 3
```{r}
baseline = read_csv('.\\data_mci\\MCI_baseline.csv',skip = 1)|>
  janitor::clean_names()|>
  mutate(
    sex = as.character(sex),
    apoe4 = as.character(apoe4))|>
  filter((age_at_onset=='.') |(age_at_onset>current_age))
```
When I imported the dataset, I skipped the first row,which is a detailed explanation of the variables.I also use `clean_names` method to lowercase the variables. I transferred variable `apoe4` and `sex` to character form because they represent a state or category.
Then I removed participants who had MCI at baseline. `r nrow(baseline)` participants are recruited and `r nrow(filter(baseline,age_at_onset !='.'))` develop MCI.The average baseline age is `r mean(pull(baseline,current_age))`.The proportion of women in the study are APOE4 carriers is `r nrow(filter(baseline,sex==0 & apoe4 =='1'))/nrow(filter(baseline,sex==0))`.


Then,import, clean and tidy the dataset of biomarker.

```{r}
amyloid = read_csv('.\\data_mci\\mci_amyloid.csv',
                   col_types = cols(.default = "d"),skip=1)|>
  janitor::clean_names()

```
When importing the amyloid dataset, I assigned the types of columns to be double,and skipped the first row which is the explanation of variables.Then I cleaned the names of variables.
